Select-based random access \\ to variable-byte encodings

### 2020-05-24 gradu/msc_english_contents.tex
Adds Haswell machine word text

The $rank$ operation can be created with a small number of instructions utilizing a modern \textsc{POPCOUNT} machine instruction \citep{Zho13}. The $select$ approach is 
known to be slower, ranging from 50 to 80 machine instructions. Intel's Haswell CPUs offer two new interesting machine instructions, \textsc{PDEP} and \textsc{TZCNT}. \textsc{PDEP} stores bits
from its first operand based on the location of 1 bits in the second operand. \textsc{TZCNT} is a machine instruction to count trailing zeros. These two can be used to create a $select_1$ query on 
a machine word with just 4 instructions \citep{Pan17a}. They have been tested with the $select$ of the SDSL-library and show an improvement of 2-15\% \citep{Pan17b}. Additionally, the 
$select$ implementation proposed in this thesis uses a trailing zero count to calculate the length of the element. Some improvement in speed may happen if the implementation is changed 
to use \textsc{TZCNT} instead.

>>>>

The $rank$ operation can be created with a small number of instructions utilizing a modern \textsc{POPCOUNT} machine instruction \citep{Zho13}. The $select$ approach is 
known to be slower, ranging from 50 to 80 machine instructions. Intel's Haswell CPUs offer two new interesting machine instructions, \textsc{PDEP} and \textsc{TZCNT}. \textsc{PDEP} stores bits
from its first operand based on the location of 1 bits in the second operand. \textsc{TZCNT} is a machine instruction to count trailing zeros. These two can be used to create a $select_1$ query on 
a machine word with just 4 instructions \citep{Pan17a}. They have been tested with the $select$ algorithm of the SDSL-library (which is also used in the proposed method of this thesis) 
and show an improvement of 2-15\% \citep{Pan17b}. Additionally, a trailing zero count function is used to calculate the length of the element. 
Some improvement in speed may happen if the implementation is changed to use \textsc{TZCNT} machine instruction instead.
